TrenchR: an R package for modular and accessbile microclimate and biophysical ecology

### 2022-10-31 Manuscript_AppS1.Rmd
Ms updates

# Assumptions
tau <- 0.7 # Atmospheric transmissivity # MRK change from 0.7 to 0.75
rho <- 0.6 # Albedo
Tb0 <- 25  # Initial assumption of body temperature (C)
doy <- day_of_year("2021-06-15", format = "%Y-%m-%d") # Day of year, changed to mid June to correspond to micro_global data
z <- 0.02 # height of animal
psa_g <- 0.33 # fraction of animal surface contacting the ground
K_sub <- 0.5 # parameter for thermal conductivity of substrate, W/mC
```

@@

We then Estimate and compare conduction.
```{r}
Qcond <- rep(NA, 24)
Qcond_NMR <- Qcond

for(hr in 1:24) {
  Qcond[hr] <- Qconduction_animal(T_g = Ts[hr] + 273.15, T_b = Ta_liz[hr] + 273.15, d = 0.025, K = K_sub, A = A, proportion = psa_g)

>>>>

# Assumptions
tau <- 0.7 # Atmospheric transmissivity # MRK change from 0.7 to 0.75
rho <- 0.6 # Albedo
Tb0 <- 25  # Initial assumption of body temperature (C)
doy <- day_of_year("2021-06-15", format = "%Y-%m-%d") # Day of year, changed to mid June to correspond to micro_global data
z <- 0.02 # height of animal
psa_g <- 0.33 # fraction of animal surface contacting the ground
K_skin <- 0.5 # parameter for thermal conductivity of skin, W/mC
```

@@

We then estimate and compare conduction.
```{r}
Qcond <- rep(NA, 24)
Qcond_NMR <- Qcond

for(hr in 1:24) {
  Qcond[hr] <- Qconduction_animal(T_g = Ts[hr] + 273.15, T_b = Ta_liz[hr] + 273.15, d = 0.025, K = K_skin, A = A, proportion = psa_g)

### 2022-10-31 Manuscript_full.Rmd
Ms updates

---
title: 'TrenchR: an R package for modular and accessbile microclimate and biophysical ecology'
author: 
  - "Lauren B. Buckley, Bryan A. Briones Ortiz, Isaac Caruso, Aji John, Ofir Levy,"
  - "Abigail V. Meyer, Eric A. Riddell, Yutaro Sakairi, and Juniper Simonis"
output:
  html_document: default
  word_document: default
  pdf_document: 
    pandoc_args: --listings
    includes:
      in_header: preamble.tex
bibliography: bibliography.bib
csl: plos.csl
editor_options: 
  markdown: 
    wrap: 72
---

@@

Few of the functions in TrenchR are "state of the art". The development of increasingly complex microclimate and biophysical models before open and reproducible science was emphasized or feasible has limited their uptake. Yet, such models are crucial to improving understand and prediction of species responses to climate change. We contend there is a need for simple and modular functions that can be combined and extended in a null modeling approach until the model adequately describes organismal interactions with their environment while remaining accessible and transparent.

@@

  - *radiative heat exchanges* of solar and thermal radiation; 
  - *convective heat exchanges* between organisms and their surrounding fluid (air or water) driven by fluid flow; 
  - *conductive heat exchanges* between organisms and solid surfaces (generally the ground) due to physical contact; and 
  - *evaporative heat exchanges* associated with organisms' evaporative water loss.

@@

where $Q_{net}$ is the net energy exchange with the environment (W), $Q_{abs}$ is the solar radiation absorbed (W), $Q_{emit}$ is the net thermal radiation emitted (W), $Q_{conv}$ is energy exchange due to convection (W), $Q_{cond}$ is energy exchange due to conduction (W), $Q_{met}$ is the energy generated by metabolism (W), and $Q_{evap}$ is the energy generated by evaporative water loss (W). 
We will estimate each term on the right side of the equation in turn. 
Estimating $Q_{abs}$ requires the surface area exposed to radiation and the solar absorptivity of the animal surface ($a$ proportion). We use zenith angle $psi$ to estimate the projected (silhouette) area as a portion of the surface area of the organism, which allows estimating absorbed solar radiation.
We model a 10 gram *Sceloporus* lizard with solar absorptivity $a = 0.9$ [@gates1980biophysical]. 
We will initially assume $T_b = T_a$ to illustrate the calculations before solving for $T_b$ given the environmental conditions.

@@

We next estimate convection $Q_{conv}$ (W) and conduction $Q_{cond}$ (W). 
We will estimate the lizard's heat transfer coefficient, $H_L$ ($Wm^{-2}K^{-1}$) using an empirical relationship for lizards (`heat_transfer_coefficient()`). We average thermal conductivity and kinematic viscosity across the day for simplicity and since there is not substantial diurnal variation. We also illustrate a function estimating $H_L$ using a spherical approximation (`heat_transfer_coefficient_approximation()`) and a simplified approximation (`heat_transfer_coefficient_simple()`) for cases when taxon specific relationships for estimating heat transfer coefficients are not available. We estimate the characteristic dimension, which determines exposure to convective heat exchange as the the cube root of volume, assuming the animal density approximates that of water [@mitchell1976heat]. These coefficients assume convection is forced by the wind. TrenchR includes approaches for free convection and a function (`free_or_forced_convection()`) that can be used to decide whether free or forced convection is appropriate. The function uses dimensionless numbers, which have been developed to describe heat transfer coefficients associated with convection over different geometries and can be estimated using TrenchR (e.g., Grashof, Nusselt, and Reynolds numbers).

```{r}
# Use DRYAIR from NicheMapR to estimate the thermal conductivity of air and kinematic viscosity

@@

where an enhancement factor, $ef$, multiplier can be incorporated to account for increases in heat exchange resulting from air turbulence in field conditions. 
We implement the function in R assuming that 2/3 of the lizard's surface area is exchanging heat through convection. 
We currently estimate no convection since we are assuming $T_b = T_a$ for illustration purposes:

@@

  Qconv[hour] <- Qconvection(T_a = Ta_liz[hour] + 273.15, T_b = Ta_liz[hour] + 273.15, H = H_L, A = A, proportion = 0.67, ef = 1.3)

@@

$$Q_{cond} = A \cdot \mbox{proportion} \cdot K(T_g-T_b)/d. $$

We implement the estimate assuming that conductive heat exchange occurs down to a soil depth of 2.5cm. We use this value rather than skin thickness, which results in rapid conduction and does not readily reach steady state conditions.

@@

  Qcond[hr] <- Qconduction_animal(T_g = Ts[hr] + 273.15, T_b = Ta_liz[hr] + 273.15, d = 0.025, K = K, A = A, proportion = psa_g)

@@

## Authors' Contributions

LB conceived and designed the package and lead all aspect of the packages creation and dissemination. All authors wrote and edited functions and associated package resources, contributed critically to manuscript drafts, and gave final approval for publication. AJ and JS configured the package. AM, BBO, YS, and LB converted and edited the educational modules. 
Add? Conceptualization, Project Administration, Software, Writing – Original Draft Preparation, Writing – Review & Editing

>>>>

---
title: 'TrenchR: an R package for modular and accessbile microclimate and biophysical ecology'
author: 
  - "Lauren B. Buckley, Bryan A. Briones Ortiz, Isaac Caruso, Aji John, Ofir Levy,"
  - "Abigail V. Meyer, Eric A. Riddell, Yutaro Sakairi, and Juniper Simonis"
output:
  word_document: default
  html_document: default
  pdf_document: 
    pandoc_args: --listings
    includes:
      in_header: preamble.tex
bibliography: bibliography.bib
csl: plos.csl
editor_options: 
  markdown: 
    wrap: 72
---

@@

The development of increasingly complex microclimate and biophysical models before open and reproducible science was emphasized or feasible has limited their uptake. Yet, such models are crucial to improving understand and prediction of species responses to climate change. We contend there is a need for simple and modular functions that can be combined and extended in a null modeling approach until the model adequately describes organismal interactions with their environment while remaining accessible and transparent.

@@

  - *radiative heat exchanges* of solar and thermal radiation; 
  - *convective heat exchanges* between organisms and their surrounding fluid (air or water) driven by fluid flow; 
  - *conductive heat exchanges* between organisms and solid surfaces (generally the ground) due to physical contact; and 
  - *evaporative and metabolic heat exchanges* associated with organisms' evaporative water loss and metabolic heat generation.

@@

where $Q_{net}$ is the net energy exchange with the environment (W), $Q_{abs}$ is the solar radiation absorbed (W), $Q_{emit}$ is the net thermal radiation emitted (W), $Q_{conv}$ is energy exchange due to convection (W), $Q_{cond}$ is energy exchange due to conduction (W), $Q_{met}$ is the energy generated by metabolism (W), and $Q_{evap}$ is the energy generated by evaporative water loss (W). 
We will estimate each term on the right side of the equation in turn. 
Estimating $Q_{abs}$ requires the surface area exposed to radiation and the solar absorptivity of the animal surface ($a$ proportion). We use zenith angle $psi$ to estimate the projected (silhouette) area as a portion of the surface area of the organism, which allows estimating absorbed solar radiation.
We model a 10 gram *Sceloporus* lizard with solar absorptivity $a = 0.9$ [@gates1980biophysical]. 
We will initially assume $T_b = T_a + 10$ to illustrate the calculations before solving for $T_b$ given the environmental conditions.

@@

We next estimate convection $Q_{conv}$ (W) and conduction $Q_{cond}$ (W). 
We will estimate the lizard's heat transfer coefficient, $H_L$ ($Wm^{-2}K^{-1}$) using an empirical relationship for lizards (`heat_transfer_coefficient()`). We average thermal conductivity and kinematic viscosity across the day for simplicity and since there is not substantial diurnal variation. We also illustrate a function estimating $H_L$ using a spherical approximation (`heat_transfer_coefficient_approximation()`) and a simplified approximation (`heat_transfer_coefficient_simple()`) for cases when taxon specific relationships for estimating heat transfer coefficients are not available. We estimate the characteristic dimension, which determines exposure to convective heat exchange as the cube root of volume, assuming the animal density approximates that of water [@mitchell1976heat].

These coefficients assume convection is forced by the wind. TrenchR includes approaches for free convection and a function (`free_or_forced_convection()`) that can be used to decide whether free or forced convection is appropriate. The function uses dimensionless numbers, which have been developed to describe heat transfer coefficients associated with convection over different geometries and can be estimated using TrenchR (e.g., Grashof, Nusselt, and Reynolds numbers).

```{r}
# Use DRYAIR from NicheMapR to estimate the thermal conductivity of air and kinematic viscosity.

@@

where an enhancement factor, $ef$, multiplier can be incorporated to account for increases in heat exchange resulting from air turbulence in field conditions. 
We implement the function in R assuming that 2/3 of the lizard's surface area is exchanging heat through convection.

@@

  Qconv[hour] <- Qconvection(T_a = Ta_liz[hour] + 273.15, T_b = Ta_liz[hour] + 10 + 273.15, H = H_L, A = A, proportion = 0.67, ef = 1.3)

@@

$$Q_{cond} = A \cdot \mbox{proportion} \cdot K_{skin}(T_g-T_b)/d$$ 
where $K_{skin}$ in the thermal conductivity of lizard skin ($Wm^-2K^-1$). We implement the estimate assuming that conductive heat exchange occurs down to a soil depth of 2.5cm. We use this value rather than skin thickness, which results in rapid conduction and does not readily reach steady state conditions.

@@

  Qcond[hr] <- Qconduction_animal(T_g = Ts[hr] + 273.15, T_b = Ta_liz[hr] + 10 + 273.15, d = 0.025, K = 0.5, A = A, proportion = psa_g)

@@

## Authors' Contributions
Lauren Buckley: Conceptualization, Methodology, Project Administration, Writing – Original Draft Preparation. All authors: Software, Writing – Review & Editing.
